---
title: "Indication Crosswalk"
author: "Ben Berger"
date: "June 5, 2017"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rio)
library(dplyr)

data <- import('data/drugapproval_tempfile.dta')
data <- data[-nrow(data),] #remove last row
colnames(data)[colnames(data) == "indication"] <- "condition"

ind2icd9.xwalk <- import('data/Cortellis_Drug_Indication_ICD9_Crosswalk.dta')
icd92hcc.xwalk <- import('data/icd2hccxw2014.dta')
icd92ccs.xwalk <- import('data/dxicd2ccsxw.dta')

icd92ccs.xwalk <- icd92ccs.xwalk[-1, ]

```

##Goal:
Match descriptive condition categories to dataset of drug approvals. 

##Issues:
Cortellis indication categories are too fine, as are 3-digit ICD-9 groupings, and it is not obvious how we should proceed in aggregating them into coarser groupings.

##What we have:

__1. Dataset of drug approvals with Cortellis Indications.__

```{r approval.data, echo = FALSE, comment = NA}
table <- head(data)[,c("brandname", "condition")]
print.data.frame(table, row.names = FALSE)
```

__2. Crosswalk of Cortellis Indications -> 3-digit ICD-9 codes.__

```{r ind2icd9.xwalk, echo = FALSE, comment = NA}
table <- head(ind2icd9.xwalk)
print.data.frame(table, row.names = FALSE)

cat("Sorted by ICD-9 code....")
table <- head(ind2icd9.xwalk[order(ind2icd9.xwalk$icd9),]) 
print.data.frame(table, row.names = FALSE)

```

\pagebreak  
__3. ICD-9 to HCC or CCS crosswalks.__

* [ICD-9 -> HCC crosswalk from NBER](http://www.nber.org/data/icd-hcc-crosswalk-icd-rxhcc-crosswalk.html)  
The first three digits of each ICD-9 correspond to the three-digit groupings used in the Indication -> ICD-9 crosswalk. Note these 3-digit ICD-9 groupings __do not__ in general map to a single HCC. Also note that only HCCs that are included in CMS's risk-adjustment model in a particular year are present. 

```{r icd92hcc.xwalk, echo = FALSE, comment = NA}
table <- head(icd92hcc.xwalk[,c(1,2,6)])
print.data.frame(table, row.names = FALSE)

```
  
  
* [ICD-9 -> CCS crosswalk from NBER](http://www.nber.org/data/icd-ccs-crosswalk.html)  
Unlike the HCC crosswalk, the CCS crosswalk is year-invariant (ICD-9 codes map to the same HCC regardless of year) and includes a much larger set of ICD-9 codes. Three-digit ICD-9 groupings __do not__ in general map to a single CCS. CCS codes, category labels, and associated ICD-9s are listed _[here](https://www.hcup-us.ahrq.gov/toolssoftware/ccs/AppendixASingleDX.txt)_. CCS categories tend to be granular enough to distinguish between related conditions while still aggregating into useful categories (e.g. all leukemias are coded CCS 39).

```{r icd92ccs.xwalk, echo = FALSE, comment = NA}
table <- head(icd92ccs.xwalk[,c(1:4)], n = 10)
print.data.frame(table, row.names = FALSE)

```

##Summary Statistics:
```{r sumstats, echo=FALSE, comment = NA}
nInd.Xwalk <- length(ind2icd9.xwalk$condition)
nInd.data <- length(unique(data$condition))

cat("There are", nInd.Xwalk, "unique Cortellis indications in the Indication -> 3-digit ICD-9 Crosswalk.")
cat("Of these,", nInd.data, "indications are represented in the drug approval data.")

cat(sum(table(data$icd9)==1), "drug approvals map to a unique 3-digit ICD-9 code.")
cat(sum(table(data$icd9[data$breakthrough == 1]) == 1), "breakthrough drug approvals map to a unique 3-digit ICD-9 code.")

cat("On average, a 3-digit ICD-9 group contains", mean(table(data$icd9)) %>% round(digits = 2), "drugs, with a median of", median(table(data$icd9)), ".")


icd92ccs.xwalk$icd.3 <- substr(icd92ccs.xwalk$icd, 1, 3)
grouped.xwalk <- icd92ccs.xwalk %>% group_by(icd.3) %>% summarise(unique_ccs = n_distinct(ccs))

cat("On average, a 3-digit ICD-9 group corresponds to", mean(grouped.xwalk$unique_ccs) %>% round(digits = 2), "unique CCS codes.")

pct <- table(grouped.xwalk$unique_ccs)/length(grouped.xwalk$unique_ccs) %>% round(digits = 2)
barplot(pct, main = 'Number of Associated HCCs for 3-digit ICD-9 Groupings', xlab = 'Number of HCCs', ylab = 'Percent of 3-digit Groups')

```

##Recommendation:
I recommend we aggregate indications into CCS codes. Although there are over 1,900 Cortellis indications in the Indication -> 3-digit ICD-9 crosswalk, only around 300 unique indications are present in the drug approval data.  
We can either do this by  

1. Looking up the entire 4-or-5-digit ICD-9 code for the indication and mapping that to a CCS code using a crosswalk, or  
2. Looking up the CCS code alone.  
  
I am happy to work on this and don't expect it to take ridiculously long.
